% Chapter 1

\chapter{Clinical Background} % Main chapter title

\label{chapter1} % For referencing the chapter elsewhere, use \ref{Chapter1}


\section{Importance of Biomarkers in ICU Studies}

Allocation of resources to patients to minimize mortality is a constant priority for healthcare professionals. This is especially important in the area of healthcare we have chosen to focus on in this paper: critical care,  where resources such as equipment and attention of specialists are even more scarce. A critical care specialist focuses on the most vulnerable and urgent patients who are placed in an Intensive Care Unit (ICU). In this setting, the specialist is often faced with difficult decisions of which patients to allocate resources to. An unfortunate reminder that has recently put such decisions under the spotlight is the current COVID-19 global pandemic. A recent study on ICU capacity in Wuhan, the disease's epicentre in China, states that at a point during the current crisis the number of COVID-19 patients who need ICU resources is 1120 while only 600 ICU beds existed. As a result, only 25\% of the patients who had died by the time of the study received the intubation and mechanical ventilation that they required \citep{wu2020characteristics}. In Lombardy, the disease's epicentre in Italy, a study states that under pre-crisis conditions,  the city's  total ICU capacity of 720 already operates at 85\% - 90\% during winter months. To make things worse, during the first two weeks after the city's first confirmed COVID-19 case, the number of COVID-19 related ICU admissions rose exponentially to 556. Moreover, estimates for the total number of COVID-19 ICU admissions in the following two weeks suggest an even more drastic shortage with linear estimates at 869 and exponential estimates at 14,542 \citep{grasselli2020}. Hence, ICU capacities are stressed under normal conditions and even further during times of crises.  Faced with such a perpetual dilemma of resource allocation, a critical care specialist uses various biomarkers in order to try and predict severe outcomes in the patients cohort. As such, a persistent priority in medical research is the discovery and analysis of connections between various biomarkers and unfavourable patient outcomes. 

\section{Problem: PF Ratio is an Important but Challenging Biomarker}

One such biomarker that is tracked in ICU settings is the  \PF(PF ratio). It is used to monitor the patient's pulmonary functions. The numerator, \Pa refers to the partial pressure of oxygen in arterial blood. It is measured in mmHg via drawing a sample of blood from an artery in the wrist or groin, and testing it in the laboratory. The denominator, \Fi refers to the initial fraction of inspired oxygen and is approximately 21\% in breathable atmospheric air. In an ICU, it can be controlled by providing the patient with a form of oxygen therapy through to the use of devices such as mechanical ventilators. The PF ratio is most notably used in the diagnosis of extreme illnesses such as Acute Respiratory Distress Syndrome (ARDS) \citep{bernard1994american}. Moreover, it has been shown to be a significant identifier of mortality risk the general ICU population \citep{villar2011risk} and as a predictor of mortality in specific subsets of patients such as newborns with Meconium Aspiration Syndrome (MAS)  \citep{narayanan2019pao2} and post-operation cardiac surgery patients \citep{esteve2014evaluation}.

However, there is a major challenge associated with the use of \Pa - its measurement. The  procedure is invasive and delayed, making it not feasible to track \Pa at frequent intervals measure for all patients. 


\section{Motivation: Can SF Ratio be an Alternative to PF Ratio and for who?}
A more convenient biomarker to measure, however, is \Sp  or peripheral capillary oxygen saturation. It is defined as the ratio of oxygenated haemoglobin by the total amount of haemoglobin in the blood. It is measured using pulse oximetry which uses the principle that oxygenated and deoxygenated haemoglobin have different absorption spectra at particular wavelengths of light \citep{jubran1999pulse} . The oximeter illuminates light at specific wavelengths through the skin (usually at the fingertips) and almost instantaneously calculates the ratio of absorption of these wavelengths to extrapolate the proportion of oxygenated haemoglobin in the blood, or \Sp \citep{jubran2015}. Therefore, unlike \Pa, \Sp can be measured in a non-invasive and instantaneous manner. 

However, a critical difference between, \Pa and \Sp is that the former a generally more accurate measure of a patient's oxygenation level since it is measured directly from a main artery while the latter is measured at the end of capillaries. Nonetheless, recent studies have shown that \SF (SF ratio) has been shown to be a non-invasive surrogate for \PF to diagnose certain subsets of patients such as children with ALI or ARDS \citep{rice2007comparison} and children with smoke inhalation injury \citep{cambiaso2017correlation}. Moreover, a retrospective study found that the SpO\textsubscript{2}/\Fi Time-at-Risk (SF-TAR), defined as the total time spent with severe hypoxemia (SF ratio $\leq 145$), is not only significantly correlated with hospital mortality for mechanically ventilated patients, but is as good or a better predictor of it than arterial gas-derived measurements of the PF ratio. 

There have also been several studies that aim to link \Sp separately to mortality. In 2015, the Tromsø study concluded that an \Sp $\leq95\%$ is associated with all-cause mortality and mortality caused by pulmonary diseases (over a 10-year follow-up period) after adjusting for sex, age, history of smoking, self-reported diseases and respiratory symptoms, BMI, and CRP concentration. When Forced Expiratory Volume (FEV1) was included as a covariate, the correlation remained significant for mortality due to pulmonary diseases but no longer significant for all-cause mortality \citep{vold2015low}. 

However, not all studies investigating a link between \Sp or \\ \SF and mortality have yielded significant results. A prospectively planned meta-analysis study using participant data from 5 randomized clinical trials (conducted from 2005-2014) of infants born before 28 weeks' gestation period found no significant difference between a lower \Sp target range (85\%-89\%) and a higher \Sp target range (91\%-95\%) on mortality or major disability at a corrected age of 18 to 24 months \citep{askie2018association}. Therefore, it seems that the use of \Sp as a predictor of mortality might not be applicable to all patient phenotypes, with potential for further sub-phenotyping. Such differences between the subpopulation might also be expected for the SF ratio which includes \Sp in the numerator. 


\section{Research Question} 

The main goal of this capstone can be summarized as follows: \\
\textit{ Investigate whether \SF is a statistically significant predictor of mortality in general ICU patient population or subsets thereof using a retrospective analysis of ICU patient records.
}

By focusing on \SF instead of only \Sp, we also implicitly hypothesize that the former is a more helpful predictor as it allows us to account for the different levels of mechanical ventilation that an ICU patient receives. In essence, it allows us to account for the patient's ability to convert inspired oxygen to peripheral oxygen saturation at the tissue level. 


\section{Data}

For this capstone I will be using \textbf{MIMIC III}, an openly available relational database developed by the MIT Lab for Computational Physiology. It contains de-identified data of 61,532 intensive care unit stays: 53,432 stays for adult patients and 8,100 for neonatal patients at the Beth Israel Deaconess Medical Center over the June 2001 - October 2012. It includes demographics, vital signs, laboratory tests, medications, mortality, etc. The database is divided into different tables of data that contain information about a patient's stay and are linked to each via identifiers such as a unique hospital admission ID and a unique patient ID. 



This is the end of the chapter. 

